---
title: "Course Project 1 - Reproducible Research"
author: "Matheus Nogueira"
date: "8/28/2017"
output:
  html_document:
    keep_md: true
---

```{r setup, include=FALSE}
rm(list = ls())
knitr::opts_chunk$set(echo = TRUE)
setwd("~/Dropbox/COURSERA/1. REPRODUCIBLE RESEARCH/Project 1")
```

#Loading and preprocessing the data

This is the code to load the data.
```{r loading Data}
data <- read.csv("activity.csv")
summary(data)
```

#What is mean total number of steps taken per day?

Now, let's take a look at the number of steps taken per day with dplyr:

```{r steps_per_day,message=FALSE}
library(dplyr)
steps_per_day <- data %>% group_by(date) %>% summarise(steps=sum(steps))
```

The histogram of the total number of steps taken each day.

```{r hist_steps_per_day}
hist(steps_per_day$steps,main = "Histogram of the total number of steps taken each day",xlab = "Number of steps per day", breaks = 10)
```

Calculating mean and median of total number of steps taken each day.

```{r mean_and_median}
mean_SPD <- round(mean(steps_per_day$steps[!is.na(steps_per_day$steps)]),digits = 2)
median_SPD <- round(median(steps_per_day$steps[!is.na(steps_per_day$steps)]),digits = 2)

mean_SPD
median_SPD
```

The mean of steps taken per day is `r mean_SPD` and the median is `r median_SPD` 

#What is the average daily activity pattern?

Now, let's change the paradigma and take a look at the daily activity patern.
First, we calculate the average number of steps taken, averaged across all days.

```{r steps_per_daily_interval}
steps_per_intervals <- data[!is.na(data$steps),] %>% group_by(interval) %>% summarise(steps=mean(steps))
```

This is the plot 5-minute interval Vs average number of steps taken

```{r plot_steps_Vs_intervals }
plot(steps_per_intervals$interval,steps_per_intervals$steps,type = "l",xlab="Intervals",ylab="Steps",main="5-minute interval and the average number of steps taken")
```


```{r max_intervals_steps}
maxIntervalSteps <- steps_per_intervals$interval[which.max(steps_per_intervals$steps)]
maxIntervalSteps
```

The 5-minute interval, on average across all the days in the dataset, that contains the maximum number of steps is `r maxIntervalSteps`.

#Imputing missing values

```{r number_of_NAs}
NumberOfNAs <- nrow(data[is.na(data[,1]),])
NumberOfNAs
```

The number of NA observations is `r NumberOfNAs`.

The strategy for imputing missing values is the mean of the interval.

```{r imputing_NAs}
new_data <- data

for (i in which(is.na(data$steps)==TRUE)) {
        interval <- data$interval[i]
        new_data$steps[i] <- steps_per_intervals$steps[which(steps_per_intervals$interval==interval)]
}
```

Histogram, mean e median of total number of steps taken per day AFTER MISSING DATA IMPUTATION.

```{r steps_per_day_2}
new_steps_per_day <- new_data %>% group_by(date) %>% summarise(steps=sum(steps))
```

The histogram of the total number of steps taken each day AFTER MISSING DATA IMPUTATION.
```{r hist_steps_per_day_2}
hist(new_steps_per_day$steps,main = "Histogram of the total number of steps taken each day after MISSING DATA IMPUTATION",xlab = "Number of steps per day", breaks = 10)
```


```{r mean_and_median_2}
new_mean_SPD <- round(mean(new_steps_per_day$steps),digits = 2)
new_median_SPD <- round(median(new_steps_per_day$steps),digits = 2)

new_mean_SPD
new_median_SPD
```

The mean of steps taken per day after missing data imputation is `r new_mean_SPD` and the median is `r new_median_SPD`. **We see that only the median is different from the first part of this report.**

#Are there differences in activity patterns between weekdays and weekends?

Creating the variable weekday/weekend:

```{r weekday_weekend}
dayweek <- weekdays(as.Date(new_data[,2]))
sat <- which(dayweek=="Saturday")
sun <- which(dayweek=="Sunday")

weekend <- union(sat,sun)
weekday_weekend <- rep("weekday",nrow(new_data))
weekday_weekend[weekend] <- "weekend"

new_data$weekday_weekend <- weekday_weekend
```

Time series plot for weekday and weekend:

```{r steps_per_daily_interval_2}
intervals <- levels(as.factor(new_data[,3]))
spi_weekday <- rep(0,length(intervals))
spi_weekend <- rep(0,length(intervals))

data_weekday <- new_data[which(new_data[,4]=="weekday"),]
data_weekend <- new_data[which(new_data[,4]=="weekend"),]

for (i in 1:length(intervals)) {
        spi_weekday[i] <- mean(data_weekday[which(data_weekday[,3]==intervals[i]),1])
        spi_weekend[i] <- mean(data_weekend[which(data_weekend[,3]==intervals[i]),1])
}
```

```{r plot_weekday_weekend}
par(mfrow=c(2,1))
plot(intervals,spi_weekday,type="l",main = "WEEKDAY",xlab="Intervals",ylab="Steps",lwd=2)
plot(intervals,spi_weekend,type="l",main = "WEEKEND",xlab="Intervals",ylab="Steps",lwd=2)
```
